---
title: "GitLab Project Analyzer (GPA)"
output: 
  flexdashboard::flex_dashboard:
    logo: www/hex_GPA_small.png
    orientation: rows
    vertical_layout: fill
    css: styles.css
    source_code: https://gitlab.devforce.disa.mil/netc-dsd/pmo/GPA
runtime: shiny
resource_files:
  - .Renviron
---

```{r libraries, include=FALSE}
library(gitlabr)
library(janitor)
library(dplyr)
library(tidyr)
library(forcats)
library(lubridate)
library(plotly)
library(shiny)
library(shinydashboard)
library(flexdashboard)
```

```{r access_api}

# DSD - GitLab 
personal_gitlab_url <- "https://gitlab.com/"

# connect as a fixed user to a GitLab instance for the session
gitlabr::set_gitlab_connection(
  gitlab_url = personal_gitlab_url,
  private_token = Sys.getenv("GITLAB_PERSONAL_TOKEN"))

# All DevForce GitLab Projects
gl_projects <- 
  gitlabr::gl_list_projects(
    max_page = 100, 
    owner = FALSE) %>% 
  janitor::clean_names()

```

```{r prep_data}
# select, filter, format, clean raw data
#   select - remove the majority of the columns for faster processing
#   filter - out all user name spaces to leave only groups
#   format - create L1/2/3 levels to assist is bucketing
#   clean - general cleaning as required
prep_gpa <- gl_projects %>% 
  # dplyr::select(
  #   id:default_branch,
  #   forks_count:namespace_full_path,
  #   packages_enabled:visibility) %>% 
  dplyr::filter(namespace_kind == "group") %>% 
  tidyr::separate(
    col = namespace_full_path, 
    sep = "/", 
    into = c("L1", "L2", "L3"), 
    remove = FALSE, 
    convert = TRUE) %>% 
  # dplyr::mutate(
  #   L2 = dplyr::if_else(
  #     condition = is.na(L2), 
  #     true = L1, 
  #     false = L2),
  #   L3 = dplyr::if_else(
  #     condition = is.na(L3), 
  #     true = L2, 
  #     false = L3)) %>% 
  dplyr::mutate(
    last_activity_at = lubridate::as_date(last_activity_at),
    created_at = lubridate::as_date(created_at))

```

Overview
=====================================

```{r prep_overview}
# develop metrics data 
#   total projects
#   number of internal projects
#   number of private projects
#   average time a project is active
tot_projects <- nrow(prep_gpa)
num_internal <- prep_gpa %>% 
  dplyr::count(visibility) %>% 
  dplyr::filter(visibility == "internal") %>% 
  dplyr::pull(n)
num_private <- prep_gpa %>% 
  dplyr::count(visibility) %>% 
  dplyr::filter(visibility == "private") %>% 
  dplyr::pull(n)
ave_life <- prep_gpa %>% 
  dplyr::mutate(project_length = last_activity_at - created_at) %>% 
  dplyr::summarise(ave_life = mean(project_length)) %>% 
  dplyr::pull(ave_life)
med_life <- prep_gpa %>% 
  dplyr::mutate(project_length = last_activity_at - created_at) %>% 
  dplyr::summarise(med_life = median(project_length)) %>% 
  dplyr::pull(med_life)
```

The GPA v0.0.6 report is generated using the [`gitlabr`](https://github.com/jirkalewandowski/gitlabr) package accessing [DevForce GitLab Projects](https://gitlab.devforce.disa.mil/). The GPA report displays metrics only for the __master branch__ of a project, other branches will not register until they are merged onto the master. Developer: [MAJ Evan Munson](mailto:evan.l.munson.mil@army.mil).

Row 
-----------------------------------------------------------------------

### Total Projects

```{r tot_projects}
flexdashboard::valueBox(
  value = tot_projects, 
  color = "#fca172",
  icon = "ion-android-apps")
```

### Number of Internal Projects

```{r num_internal}
flexdashboard::valueBox(
  value = num_internal, 
  color = "#f98228",
  icon = "ion-clipboard")
```

### Number of Private Projects

```{r num_private}
flexdashboard::valueBox(
  value = num_private, 
  color = "#fc6b02",
  icon = "ion-briefcase")
```

### Project Length: Average (days)

```{r ave_life}
flexdashboard::valueBox(
  value = round(x = ave_life, digits = 1), 
  color = "#dd561c",
  icon = "ion-ios-analytics-outline")
```

### Project Length: Median (days)

```{r med_life}
flexdashboard::valueBox(
  value = round(x = med_life, digits = 1), 
  color = "#893101",
  icon = "ion-ios-barcode-outline")
```

Row
-----------------------------------------------------------------------

### Project Creation Through Time {data-width=700}

```{r projects_time}
# activity through time
prep_activity <- prep_gpa %>% 
  dplyr::mutate(
    day = lubridate::round_date(
      x = created_at,
      unit =  "month")) %>% 
  dplyr::group_by(day) %>% 
  dplyr::summarise(visits = n()) %>% 
  dplyr::arrange(dplyr::desc(visits))

plot_activity <- prep_activity %>% 
  ggplot2::ggplot(
    ggplot2::aes(
      x = day, 
      y = visits)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(
    se = FALSE,
    method = "loess",
    color = "#e24329") +
  ggplot2::labs(
    x = "Year", 
    y = "Created Projects") +
  ggplot2::theme_bw()

plotly::ggplotly(plot_activity)
```

### Top 15 Contributing Organizations {data-width=340}

```{r top_15}
# top 15 contributing organizations
prep_orgs <- prep_gpa %>% 
  dplyr::group_by(L1) %>% 
  dplyr::count() %>% 
  dplyr::arrange(dplyr::desc(n)) %>% 
  dplyr::rename("count" = "n") %>% 
  utils::head(15)

plot_orgs <- prep_orgs %>% 
  ggplot2::ggplot(ggplot2::aes(x = reorder(stringr::str_wrap(L1, 30), count), y = count)) +
  ggplot2::geom_bar(
    stat = "identity",
    color = "#fca326", 
    fill = "#fca326") +
  ggplot2::theme_bw() +
  ggplot2::labs(
    x = "Organizational Group", 
    y = "Number of Projects") +
  ggplot2::coord_flip()
  
plotly::ggplotly(plot_orgs)
```

Organization
=====================================

Sidebar {.sidebar}
-------------------------------------

### Organization Content

Please select which organization you would like to see the details of.

```{r select_content}

# Group Selector 
 shiny::selectInput(
   inputId = "groups", 
   label = "Group Selector", 
   choices = sort(unique(prep_gpa$L1)), 
   selected = "netc-dsd")

# Sub-Group Selector
# create filter for l2 entries
sub_group <- reactive({
  prep_gpa %>% 
    dplyr::filter(L1 %in% input$groups)
})

# display L2, based off L1 selection
observeEvent(sub_group(), {
  choices <- unique(sub_group()$L2)
  updateSelectInput(session = session,
                    inputId = "sub_groups",
                    choices = choices)
})

# Display L2 options
selectInput(inputId = "sub_groups",
            label = "Sub-Group Selector",
            choices = NULL)

```

Row
-----------------------------------------------------------------------

```{r prep_org}
# filter data frame by selected group (aka L1)
groups_filter <- reactive(as.character(input$groups))
sub_groups_filter <- reactive(as.character(input$sub_groups))


prep_org <- shiny::reactive({
  
  prep_gpa %>%
    dplyr::filter(
      L1 == groups_filter() &
        L2 == sub_groups_filter())
})


```

### Total Projects

```{r tot_projects_org}

renderValueBox({
  
  tot <- nrow(prep_org())
  
  flexdashboard::valueBox(
    value = tot, 
    color = "#fca172",
    icon = "ion-android-apps")
})

```

### Number of Internal Projects

```{r num_internal_org}

renderValueBox({
  
  num_internal_org <- prep_org() %>% 
    dplyr::count(visibility) %>% 
    dplyr::filter(visibility == "internal") %>% 
    dplyr::pull(n)
  
  flexdashboard::valueBox(
    value = num_internal_org, 
    color = "#f98228",
    icon = "ion-clipboard")
})

```

### Number of Private Projects

```{r num_private_org}

renderValueBox({
  
  num_private_org <- prep_org() %>% 
    dplyr::count(visibility) %>% 
    dplyr::filter(visibility == "private") %>% 
    dplyr::pull(n)
  
  flexdashboard::valueBox(
    value = num_private_org, 
    color = "#fc6b02",
    icon = "ion-briefcase")
})

```

### Project Length: Average (days)

```{r ave_life_org}

renderValueBox({
  
  ave_life_org <- prep_org() %>% 
    dplyr::mutate(
      project_length = last_activity_at - created_at) %>% 
    dplyr::summarise(
      ave_life = mean(project_length)) %>% 
    dplyr::pull(ave_life)
  
  flexdashboard::valueBox(
    value = round(x = ave_life_org, digits = 1), 
    color = "#dd561c",
    icon = "ion-ios-analytics-outline")
})

```

### Project Length: Median (days)

```{r med_life_org}

renderValueBox({
  
  med_life_org <- prep_org() %>% 
    dplyr::mutate(
      project_length = last_activity_at - created_at) %>% 
    dplyr::summarise(
      med_life = median(project_length)) %>% 
    dplyr::pull(med_life)
  
  flexdashboard::valueBox(
    value = round(x = med_life_org, digits = 1), 
    color = "#893101",
    icon = "ion-ios-barcode-outline")
})

```

Row
-----------------------------------------------------------------------

### GitLab Code Repositories {data-width=700}

```{r repo_time}
# view code repos through time by group (aka L1)

renderPlotly({
  
  prep_projects <- prep_org() %>% 
    dplyr::arrange(-dplyr::desc(created_at)) %>% 
    dplyr::mutate(
      proj_ord = dplyr::row_number()) %>% 
    dplyr::relocate(proj_ord, .before = id) %>% 
    dplyr::mutate(
      name = forcats::fct_reorder(name, proj_ord),
      name = forcats::fct_rev(name)) 
  
  plot_projects <- prep_projects %>% 
    ggplot2::ggplot(
      ggplot2::aes(y = name)) +
    ggplot2::geom_segment(
      ggplot2::aes(x = created_at,
                   xend = last_activity_at,
                   y = name, yend = name),
      color = "#e24329", 
      lineend = "round",
      size = 2) +
    ggplot2::geom_vline(
      ggplot2::aes(
        xintercept = as.numeric(Sys.Date())), 
      color = "red", 
      linetype = "dashed") +
    ggplot2::scale_x_date() +
    ggplot2::theme_bw() +
    ggplot2::labs(
      x = "Year", 
      y = "Project")
  
  plotly::ggplotly(plot_projects)
  
})

```

### Organization Sub-Groups {data-width=340}

```{r sub_group}

renderPlotly({

  # view sub group (aka L2) repo content given a group (aka L1)
  prep_levels <- prep_org() %>%
    dplyr::group_by(L3) %>%
    dplyr::count() %>%
    dplyr::arrange(dplyr::desc(n))

  plot_levels <- prep_levels %>%
    ggplot2::ggplot(
      ggplot2::aes(
        x = reorder(
          stringr::str_wrap(L3, 30),
          n),
        y = n)) +
    ggplot2::geom_bar(
      stat = "identity",
      color = "#fca326",
      fill = "#fca326") +
    ggplot2::theme_bw() +
    ggplot2::labs(
      x = "Organizational Group",
      y = "Number of Projects") +
    ggplot2::coord_flip()

  plotly::ggplotly(plot_levels)
  
})

```

Project
=====================================

Sidebar {.sidebar}
-------------------------------------

### Groups

Use the below drop-down menus to navigate to the GitLab Group, Sub-Group and if applicable, Sub-Sub-Group to find the Project you are interested in viewing the details of.

```{r select_project}

# L1
selectInput(inputId = "level_1",
            label = "Groups",
            choices = sort(unique(prep_gpa$L1)),
            selected = "netc-dsd")

# L2
# create filter for l2 entries
filter_level_1 <- reactive({
  prep_gpa %>% 
    dplyr::filter(L1 %in% input$level_1)
})

# display L2, based off L1 selection
observeEvent(filter_level_1(), {
  choices <- unique(filter_level_1()$L2)
  updateSelectInput(session = session,
                    inputId = "level_2",
                    choices = choices)
})

# Display L2 options
selectInput(inputId = "level_2",
            label = "Sub-Groups",
            choices = NULL)

# L3
# create filter for l3 entries
filter_level_2 <- reactive({
  prep_gpa %>% 
    dplyr::filter(L2 %in% input$level_2)
})

# display L2, based off L1 selection
observeEvent(filter_level_2(), {
  choices <- unique(filter_level_2()$L3)
  updateSelectInput(session = session,
                    inputId = "level_3",
                    choices = choices)
})

# Display L2 options
selectInput(inputId = "level_3",
            label = "Sub-Sub-Groups",
            choices = NULL)
```

-------------------------------------

### Projects

Select the Project you are interested in viewing the details of.

```{r project-picker}
# Project Picker
# create filter for l3 entries
filter_level_3 <- reactive({
  prep_gpa %>% 
    dplyr::filter(
      L1 %in% input$level_1 &
        L2 %in% input$level_2 &
        L3 %in% input$level_3)
})

# display Projects, based off L3 selection
observeEvent(filter_level_3(), {
  choices <- unique(filter_level_3()$name)
  updateSelectInput(session = session,
                    inputId = "project_name",
                    choices = choices)
})

# Display Project options
selectInput(inputId = "project_name",
            label = "Project",
            choices = NULL)

```

Row
-----------------------------------------------------------------------

```{r prep_project}
# filter data frame by selected group (aka L1)
filter_level_1_rec <- reactive(as.character(input$level_1))
filter_level_2_rec <- reactive(as.character(input$level_2))
filter_level_3_rec <- reactive(as.character(input$level_3))
filter_project_name_rec <- reactive(as.character(input$project_name))

prep_project <- shiny::reactive({
  
  prep_gpa %>%
    dplyr::select(
      id,
      name,
      L1,
      L2,
      L3,
      forks_count) %>% 
    dplyr::filter(
      L1 == filter_level_1_rec() &
        L2 == filter_level_2_rec() &
        L3 == filter_level_3_rec() &
        name == filter_project_name_rec())
})

# pull number of commits
project_commits <- shiny::reactive({
  
  gitlabr::gl_get_commits(project = prep_project()$id)
})

# pull number of branches
project_branches <- shiny::reactive({
  
  gitlabr::gl_list_branches(project = prep_project()$id)
})

# Mean time between commits (commit_diffs has been converted to days)
project_timedif <- shiny::reactive({
  
  project_commits() %>% 
    dplyr::select(committed_date) %>% 
    dplyr::mutate(
      committed_date = lubridate::ymd_hms(
        committed_date, 
        tz = "US/Arizona")) %>% 
    dplyr::arrange(dplyr::desc(committed_date)) %>% 
    dplyr::mutate(
      commit_diff = committed_date - dplyr::lead(committed_date),
      commit_diff = round(
        x = as.numeric(commit_diff) / 86400, 
        digits = 3))
})


```

### Number of Commits

```{r num_branches}

renderValueBox({
  
  flexdashboard::valueBox(
    value = nrow(project_commits()), 
    color = "#fca172",
    icon = "ion-network")
})

```

### Number of Branches

```{r num_commits}

renderValueBox({

  flexdashboard::valueBox(
    value = nrow(project_branches()), 
    color = "#f98228",
    icon = "ion-ios-shuffle-strong")
})

```

### Number of Forks

```{r time_commits}

renderValueBox({
  
  fork <- prep_project()$forks_count[1]

  flexdashboard::valueBox(
    value = fork, 
    color = "#fc6b02",
    icon = "ion-fork-repo")
})

```

### Commits per Day: Average (days)

```{r commits_day}

renderValueBox({

  cpd <- nrow(project_commits()) / 
    as.numeric(head(project_timedif()$committed_date, n = 1) -
                 tail(project_timedif()$committed_date, n = 1))
  
  flexdashboard::valueBox(
    value = round(x = cpd, digits = 2), 
    color = "#dd561c",
    icon = "ion-ios-analytics-outline")
})

```

### Project Length: (days)

```{r project_length}

renderValueBox({
  
  project_length <- as.numeric(head(project_timedif()$committed_date, n = 1) - 
                                 tail(project_timedif()$committed_date, n = 1))

  flexdashboard::valueBox(
    value = round(x = project_length, digits = 2), 
    color = "#893101",
    icon = "ion-ios-barcode-outline")
})

```

Row
-----------------------------------------------------------------------

### Project Commits {data-width=700}

```{r project_time}
# view code repos through time by group (aka L1)

renderPlotly({
  
  # show project activity over time
  project_activity <- project_timedif() %>% 
    dplyr::mutate(
      day = lubridate::round_date(
        x = committed_date,
        unit =  "day")) %>% 
    dplyr::group_by(day) %>%
    dplyr::summarise(commits = n())
  
  plot_activity <- project_activity %>% 
    ggplot2::ggplot(
      ggplot2::aes(
        x = day, 
        y = commits)) +
    ggplot2::geom_point() +
    ggplot2::geom_line(color = "#e24329") +
    # ggplot2::geom_smooth(
    #   se = FALSE,
    #   method = "loess",
    #   color = "#e24329") +
    ggplot2::labs(
      x = "Year", 
      y = "Created Projects") +
    ggplot2::theme_bw()
  
  plotly::ggplotly(plot_activity)
  
})

```

### Commits by Person {data-width=340}

```{r commits_person}

renderPlotly({
  
  # show the number of commits by author(s)
  project_committers <- project_commits() %>% 
    dplyr::group_by(author_name) %>% 
    dplyr::count() %>% 
    dplyr::arrange(dplyr::desc(n))
  
  committers <- project_committers %>% 
    ggplot2::ggplot(
      ggplot2::aes(
        x = reorder(
          stringr::str_wrap(author_name, 30), 
          n), 
        y = n)) +
    ggplot2::geom_bar(
      stat = "identity", 
      color = "#fca326", 
      fill = "#fca326") +
    ggplot2::theme_bw() +
    ggplot2::labs(
      title = "Top Contributiong Organizational Group", 
      x = "Organizational Group", 
      y = "Number of Projects") +
    ggplot2::coord_flip()
  
  plotly::ggplotly(committers)
  
})

```
















